Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-09-21 vesicle_code/docs/permeableVes/main.tex
added some new stenosis runs

\begin{figure}[htp]
  \centering
%  \ifTikz
  \input{figures/ellipseArea.tikz}
%  \fi
%  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The time axis is scaled by the permeability rate $\beta$. Since
  the curves collapse to a single curve, the permeability rate $\beta$
  sets a relaxation time scale.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. Depending on its reduced
area and viscosity contrast, two-dimensional vesicles either tank tread
or tumble~\cite{fin-lam-sei-gom2008,kra-win-sei-lip1996}. Since we are
considering semi-permeable vesicles, we assume there is no viscosity
contrast, and tumbling is not possible.
\begin{figure}[htp]
  \centering
  \input{figures/shearSS.tikz}
  \caption{\label{fig:shearSS} The reduced area of a semi-permeable
  vesicle in a shear flow. The final reduced area is nearly independent
  of the initial reduced area.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
Inspired by an experiment that measures the effects of cells on 
dynamical pressure-drop variations along a micrometer-sized
channel~\cite{abk-fai-sto2006}, we consider a numerical experiment where
a semi-permeable vesicle passes through a constriction. We also measure
the pressure-drop variation by measuring the pressure at the inlet and
outlet of the constriction.

In each experiment, we initialize an elliptical vesicle with longest
diameter of 9$\mu$m. We use a length scale of $R_0 = 10^{-6}$m, a
bending stiffness of $k_b = 10^{-19}$J, and fluid viscosity of $\mu=5
\times 10^{-2}$kg/ms. Using the relaxation time scale $\tau = \mu
R_0^3/k_b = 5 \times 10^{-1}$s, the pressure scale is $P = k_b/R_0^3 =
10^{-1}$Pa. The geometry is similar to an
experiment~\cite{abk-fai-sto2006} where the membrane starts in a channel
with diameter $25\mu$m and passes into a channel with length $45\mu$m
and diameter $5\mu$m (Figure~\ref{fig:stenosisGeom}. We impose a
Poiseuille flow at an inlet and outlet and a no-slip boundary condition
on the upper and lower walls. The flow rate is chosen so that the
residency time of the vesicle in the constriction agrees with the
experimental results.

>>>>

\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/ellipseArea.tikz}
  \fi
%  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The time axis is scaled by the permeability rate $\beta$. Since
  the curves collapse to a single curve, the permeability rate $\beta$
  sets a relaxation time scale.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. Depending on its reduced
area and viscosity contrast, two-dimensional vesicles either tank tread
or tumble~\cite{fin-lam-sei-gom2008,kra-win-sei-lip1996}. Since we are
considering semi-permeable vesicles, we assume there is no viscosity
contrast, and tumbling is not possible.
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/shearSS.tikz}
  \fi
  \caption{\label{fig:shearSS} The reduced area of a semi-permeable
  vesicle in a shear flow. The final reduced area is nearly independent
  of the initial reduced area.}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
We consider the effect of stenosis on a semi-permeable vesicle. To allow
for large amounts of fluid flux through the membrane, we consider
geometry with a long constricted region. We start by considering a
vesicle initialized with reduced area $\nu = 0.65$, and we compare the
vesicle shape when $\beta = 0$ (impermeable) and $\beta = 10^{-3}$
(permeable). Figure~\ref{fig:stenosisShape} shows the vesicles' shape at
several locations along the constriction, and the corresponding tensions
are in Figure~\ref{fig:stenosisTension}.
\begin{figure}[htp]
  \scalebox{0.4}{\input{figures/stenosisShape1.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape2.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape3.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape4.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape5.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape6.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape7.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape8.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape9.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape10.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape11.tikz}}
  \scalebox{0.4}{\input{figures/stenosisShape12.tikz}}
%%  \scalebox{0.5}{\input{figures/stenosisShape13.tikz}}
  \caption{\label{fig:stenosisShape} An impermeable (red) and
  semi-permeable (dashed blue) vesicle in a stenosed geometry (black).
  The vesicles' initial configuration are identical with reduced area
  $\nu = 0.65$. The semi-permeable vesicle deflates throughout the
  constriction (see Figure~\ref{fig:stenosisArea}) and develops higher
  curvature regions than its impermeable counterpart.}
\end{figure}

\begin{figure}[htp]
  \centering
  \begin{tabular}{ccc}
    \scalebox{0.31}{\input{figures/stenosisTension1.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisTension3.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisTension5.tikz}} \\
    \scalebox{0.31}{\input{figures/stenosisTension7.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisTension9.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisTension11.tikz}} 
  \end{tabular}
  \caption{\label{fig:stenosisTension} The tension of the impermeable
  (red) and semi-permeable (dashed blue) vesicles in
  Figure~\ref{fig:stenosisShape}. The point $\theta=0$ corresponds to
  the right-most point of the vesicle.}
\end{figure}

\begin{figure}[htp]
  \centering
  \begin{tabular}{ccc}
    \scalebox{0.31}{\input{figures/stenosisFlux1.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisFlux3.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisFlux5.tikz}} \\
    \scalebox{0.31}{\input{figures/stenosisFlux7.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisFlux9.tikz}} &
    \scalebox{0.31}{\input{figures/stenosisFlux11.tikz}} 
  \end{tabular}
  \caption{\label{fig:stenosisFlux} The flux (equivalent to pressure
  jump) of the impermeable (red) and semi-permeable (dashed blue)
  vesicles in Figure~\ref{fig:stenosisShape}. The point $\theta=0$
  corresponds to the right-most point of the vesicle.}
\end{figure}

\begin{figure}[htp]
  \centering
%  \ifTikz
  \input{figures/stenosisArea.tikz}
%  \fi
%  \includegraphics[width=0.8\linewidth]{figures/ellipseArea}
  \caption{\label{fig:stenosisArea} The area of a stenosed
  semi-permeable vesicle, with permeability rate $\beta=10^{-3}$. The
  $x$-axis is the $x$-coordinate of the center of mass of the vesicle.
  The vesicle loses fluid at a constant rate throughout the constriction
  that is between $\pm 60$. The semi-permeable vesicle loses 12.8\% of
  its fluid.}
\end{figure}

Inspired by an experiment that measures the effects of cells on 
dynamical pressure-drop variations along a micrometer-sized
channel~\cite{abk-fai-sto2006}, we consider a numerical experiment where
a semi-permeable vesicle passes through a constriction. We also measure
the pressure-drop variation by measuring the pressure at the inlet and
outlet of the constriction.

In each experiment, we initialize an elliptical vesicle with longest
diameter of 9$\mu$m. We use a length scale of $R_0 = 10^{-6}$m, a
bending stiffness of $k_b = 10^{-19}$J, and fluid viscosity of $\mu=5
\times 10^{-2}$kg/ms. Using the relaxation time scale $\tau = \mu
R_0^3/k_b = 5 \times 10^{-1}$s, the pressure scale is $P = k_b/R_0^3 =
10^{-1}$Pa. The geometry is similar to an
experiment~\cite{abk-fai-sto2006} where the membrane starts in a channel
with diameter $25\mu$m and passes into a channel with length $45\mu$m
and diameter $5\mu$m (Figure~\_ref{fig:stenosisGeom}. We impose a
Poiseuille flow at an inlet and outlet and a no-slip boundary condition
on the upper and lower walls. The flow rate is chosen so that the
residency time of the vesicle in the constriction agrees with the
experimental results.
